---
title: "Generating Figures for Moulana et al."
author: "Alief Moulana"
date: "7/10/2019"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

#load all necessary packages
library(reshape2)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(dplyr)
library(stringr)
library(factoextra)

#load all the data needed

pangenome <- as.matrix(read.table('similarity_matrix.txt', 
                                  quote = "",sep='\t',header=FALSE)) 
#load the similarity matrix data

cluster<- read.table('Sulfurovum_gene_clusters_summary_cleaned.txt',sep='\t',quote="",
                     header=TRUE) 
#load the summarized sulfurovum pangenome profile from anvi'o

```
## Taxonomy distribution and pangenomic profile
We performed the common metagenomic pipeline of assembly, mapping, binning, and annotation (see Methods). Then, we generated Supplementary Figure 1 on a spread sheet using the taxonomy data we obtained from "phylosift". Because Sulfurovum was the most abundant genus in both Mid Cayman Rise and Axial genomes, we used Sulfurovum MAGs in the subsequent pangenome analyses following "anvi'o" pangenome workflow. Then we captured Figure 1A using command `anvi-display-pan` with genomes storage and pangenome databases generated from the previous steps. Moreover, we obtained the summary of the pangenome profile (the cluster file) by running the command `anvi-summarize` for the PAN DB. We also analyzed how the pattern of total and core genome accumulation as more MAGs considered. We visualize the result from the chunk below (Supplementary Figure 2).

```{r pangenome.count}
cluster.basic.count <- unique(data.frame(Group=cluster$gene_cluster_id,
                                  Number=cluster$num_genomes_gene_cluster_has_hits,
                                  Genome=cluster$genome_name)) 
#create df, remove duplicates
cluster.basic.count$Group <- as.character(cluster.basic.count$Group)

#reshape df
cluster.basic.reshaped <- reshape(cluster.basic.count,v.names="Number",
                                  timevar="Genome",idvar=c("Group"),direction="wide")
cluster.basic.matrix <- as.matrix(cluster.basic.reshaped[,2:23]) #turn into matrix
cluster.basic.matrix[] <- c(TRUE,FALSE)[(is.na(cluster.basic.matrix) )+ 1] #binary

total.each.genome <- apply(cluster.basic.matrix,2,sum) #numb of gene groups per genome

#create a function returning whether or not the sum of a row up until i equals i
count.sum <- function(row,i){
  sum(row[1:i]) == i
}

#create a vector that count the number of core genes taking into account i genomes
core.now <- c()
for(j in 1:22){
  core.now[j] <- sum(apply(cluster.basic.matrix,1,count.sum,j))
}

#create a function to perform the analysis below
count.sum.2 <- function(numb,x){
  var <- sum(apply(as.matrix(cluster.basic.matrix[,1:x-1]),1,sum) ==0 & 
               cluster.basic.matrix[,x] == 1)
  return(numb+var)
}

#create a vector that expands number of genes as more genomes sequenced
total.now <- c()
total.now[1] <- total.each.genome[1]
count <- total.now[1]
for(k in 2:22){
  total.now[k] <- count.sum.2(count,k)
  count <- total.now[k]
}
#merge the core now and total now
count.groups <- data.frame(Number=c(1:22),Core=core.now,Total=total.now)
ggplot(count.groups,aes(x=Number))+
  geom_line(aes(y=Core,colour="Number of Core Genes"))+
  geom_line(aes(y=Total/5,colour="Number of Total Genes"))+
  scale_y_continuous(sec.axis = sec_axis(~.*5, name = "Number of Total Genes"))+
  xlab("Number of MAGs Recovered") +
  ylab("Number of Core Genes") +
  labs(colour = "Parameter")+
  theme_bw()+
  theme(axis.text=element_text(size=12),
        legend.text=element_text(size=12),
        legend.title=element_blank(),
        axis.title=element_text(size=14),
        legend.position = c(0.3, 0.85))
```

Moreover, we also analyze the similarity in gene content among MAGs. We first run `similarity_matrix.py` which takes the cluster file as an input and produces the file `similarity_matrix.txt`. The program produces a 22-by-22 matrix, where each entry $ij$ represents the proportion of genes in the $i$th MAG that is also contained by the $j$th MAG. We then visualize the matrix and find the distance among the genomes in the following chunk.

```{r}

```
